Exploring probability distributions for bivariate temporal granularities

### 2019-08-12 main.Rmd
defining calendar categorisation first and then move on to linear or circular

# Time granularities{#sec:algorithm}

Often we partition time into years, months, weeks, days and so on in a hierarchical manner to relate it to data. Such discrete abstractions of time can be thought of as linear time granularities [@aigner2011visualization]. Providing a formalism to this abstraction is important to model a time series across differently grained temporal domains.

@@

<!-- Representation with tsibble -->
Suppose we have a tsibble with a time index in one column and keys and variables in other columns. A time domain, as defined by Bettini, is essentially a mapping of row numbers (the index set) to the time index. A linear granularity is a mapping of row numbers to subsets of the time domain. For example, if the time index is days, then a linear granularity might be weeks, months or years.

<!-- What is calendar categorisation -->
What we need to add to this are additional categorizations of time that are not linear granularities and are useful to represent periodicity. Examples include day-of-week, time-of-day, week-of-year, day-of-month, month-of-year, workingday/non-workingday, etc. Many of these are circular, such as day-of-week, time-of-day. Some are nearly circular such as day-of-month. Some are irregular such as workingday/non-working day. Let's call all of these "calendar categorizations". Anything that maps a time index to a categorical variable can be considered a **calendar categorization**.

<!-- What is circular granularity and some examples -->
We specify the circular categorizations using modular arithmetic and call these **circular granularities**. The number of categories is essentially the periodicity of a circular time granularity. For example, suppose the time index is in minutes, and let $n_i$ be the number of categories created by the circular granularity $C_i$. Then the following categorizations can be computed.

>>>>

# Formal conceptualisation of calendar categorisations{#cirgran-def}

<!-- general calendar categorisation -->
Often we partition time into months, weeks, days and so on in a hierarchical manner to relate it to data. Such discrete abstractions of time can be thought of as time granularities [@aigner2011visualization]. Providing a formalism to this abstraction is important to model a time series across differently grained temporal domains. Examples of time abstractions may also include day-of-week, time-of-day, week-of-year, day-of-month, month-of-year, workingday/non-workingday, etc which are useful to represent periodicity in the data. Many of these are circular, such as day-of-week, time-of-day. Some are nearly circular such as day-of-month. Some are irregular such as workingday/non-working day. Let us call all of these different abstractions of time as "calendar categorizations".

These calendar categorisations can be linear, circular or aperiodic. We call the calendar categorisations as "linear" if they respect the linear progression of time. We call these _*linear time granularities*_. Examples include hours, days, weeks and months. _*Circular time granularities*_ accommodate periodicities in time such as hour of the day, and day of the week. _*Aperiodic circular granularities*_ are neither linear nor circular, such as day of the month or public holidays.

<!-- <!-- Representation with tsibble -->
<!-- Suppose we have a tsibble with a time index in one column and keys and variables in other columns. A time domain, as defined by [@Bettini1998-ed], is essentially a mapping of row numbers (the index set) to the time index. A linear granularity is a mapping of row numbers to subsets of the time domain. For example, if the time index is days, then a linear granularity might be weeks, months or years. -->

<!-- <!-- What is calendar categorisation --> 
<!-- What we need to add to this are additional categorizations of time that are not linear granularities and are useful to represent periodicity. Examples include day-of-week, time-of-day, week-of-year, day-of-month, month-of-year, workingday/non-workingday, etc. Many of these are circular, such as day-of-week, time-of-day. Some are nearly circular such as day-of-month. Some are irregular such as workingday/non-working day. Let's call all of these "calendar categorizations". Anything that maps a time index to a categorical variable can be considered a **calendar categorization**.  -->

<!-- What is circular granularity and some examples -->
<!-- We specify the circular categorizations using modular arithmetic and call these **circular granularities**. The number of categories is essentially the periodicity of a circular time granularity. For example, suppose the time index is in minutes, and let $n_i$ be the number of categories created by the circular granularity $C_i$. Then the following categorizations can be computed. -->

@@
